Reasoning about a Capability Machine with Local Capabilities\\ Provably Safe Stack and Return Pointer Management (without OS Support)\\ Technical Appendix Including Proofs and Details

### 2017-09-06 fully-abstract-comp/technical-report.tex
Added answer to Dominique. Added approximate timestamps for my comments.

\begin{document}
\section{The two capability machines}
\subsection{Domains}
Source language domains:
\[
  \begin{array}{rrcl}
   \addrbnf,\basebnf \in & \Addr & \defeq & \nats \\
    \sealbasebnf, \sigma \in & \Seal & \defeq & \nats \\
    &\Word & \defeq & \ints \uplus \SealableCaps\\
    \permbnf \in& \Perm & \defbnf & \dots \\
    &\glbnf & \defbnf & \glob \mid \local \\
    &\linbnf & \defbnf & \linear \mid \normal \\
    &\aendbnf & \in & \Addr \uplus \{\infty \} \\
    &\sealendbnf & \in & \Seal \uplus \{\infty \} \\
    &\scbnf &\defbnf & ((\permbnf,\glbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\glbnf,\sealbasebnf,\sealendbnf}\\
    & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
    &\cbnf & \defbnf &  \scbnf \mid \sealed{\scbnf,\sigma}{\sourcecolor{} \; \mid \retptrd \mid \retptrc}\\ 
    &\RegName & \defbnf &\pcreg \mid \rret \mid \rstk \mid \rdata \mid \dots \\
    &\Reg & \defeq & \RegName \fun \Word\\
    &\Mem & \defeq & \Addr \fun \Word \\
    &\MemSeg & \defeq & \Addr \parfun \Word \\
    {\sourcecolor{} \sstkframe \in} & {\sourcecolor{} \StkFrame} & {\sourcecolor{} \defeq} & {\sourcecolor{} \Caps \times \MemSeg \times \Addr}\\
    \src{\sstk \in}& \src{ \Stack} & \src{ \defeq} & \src{ \StkFrame^*} \\
    \Phi \in & \ExecConf & \defeq & \Mem \times \Reg {\sourcecolor{} \; \times \; \Stack \times \MemSeg }\\
    &\Conf & \defeq & \ExecConf \uplus \{\failed\} \uplus (\{\halted\} \times \Mem)
  \end{array}
\]
\lau{Should seals also have the option to be linear?}
The target language domains are all the non blue parts in the above. The source language domains are the black and blue parts in the above. Further
\begin{itemize}
\item $\glbnf$ defines domain $\Global$
\item $\linbnf$ defines domain $\Linear$
\item $\scbnf$ defines domain $\SealableCaps$
\item $\cbnf$ defines domain $\Caps$
\item $\Perm$ is defined as the set of permissions in Figure~\ref{fig:perm-hier}.
\end{itemize}

@@

The source machine is also a capability machine with memory capabilities, local capabilities and sealed capabilities. Unlike the target machine, the source machine is going to have a built in stack along with special stack and return capabilities. The syntax of the source machine language is as follows:
\[
  \begin{array}{rcl}
    n & \in & \nats \\
    \src{r} &\in &  \RegName \setminus \{\rstk\}\\
    \src{\rn} &\defbnf & \src{r} \mid n \\
    \src{i} & \defbnf &  \trg{i} \mid \scall{\src{r}}{\src{r}}
  \end{array}
\]
\lau{We had $\rret$ as not being one of the registers on the source machine. I could not recall why we had this, so I removed it. (I think it may have been from when we had call and return instructions?)}
There is one syntactic difference between the source language and the target language, namely the target language has an extra instruction in the $\scall{}{}$ instruction.

@@

  \caption{Permission hierarchy}
  \label{fig:perm-hier}
\end{figure}
\lau{We should maybe add a picture with local/global hieararchy.}
\lau{For now, I will add the linear capabilities such that a linear capability stays linear. It does, however, seem like it would be okay to have linear capabilities become non-linear (I at least don't see anything that would break down completely). One draw back would be that one could ``ruin'' the stack capability by making it non-linear.}

@@

\textbf{Call}\\
Call has three cases, a successful call, undefined and failed. For a call to be successful, it needs two sealed capabilities - a data and a code capability. The code capability must be non-executable and the stack register must contain a stack pointer. Futher the code and data capability must be local and there must be no attempt at passing on the stack capability.
\lau{Add more description here.}

For the call to be undefined, the conditions necessary for a legal call must be met and there must be an attempt at passing the stack pointer to the callee or the called capabilities must be non-local \lau{I have forgotten why this is an issue, but I think it is because we are crossing a security domain?}

@@

Unlike CHERI, we want sealed capabilies to be useable in a somewhat normal jump as well as the $\scall{}{}$. We cannot just use $\sjmp{}$ because it only takes one argument and we need to be able to jump to a code/data pair. To this end we introduce $\sxjmp{}{}$:
\begin{align*}
  \ssem{\sxjmp{r_1}{r_2}} = & 
                              \begin{cases}
                                \Phi'
                                \arraycolsep=0pt
                                \begin{array}[t]{l}
                                  [\reg.\pcreg \mapsto c_1] \\
                                  [\reg.\rdata \mapsto c_2]
                                \end{array} & 
                                \arraycolsep=0pt
                                \begin{array}[t]{l}
                                  \Phi(r_1) = \sealed{c_1,\sigma} \text{ and } \\
                                  \Phi(r_2) = \sealed{c_2,\sigma} \text{ and } \\
                                  \text{ for $i \in \{1,2\}$ if $\isLinear{c_i}$, then $\Phi'(r_i) = 0$} \\
                                  \text{ otherwise $\Phi'(r_i) = \Phi(r_i)$ and} \\
                                  \text{ $ \Phi(r) = \Phi'(r) \text{ for } r\in \tRegName \setminus \{r_1,r_2\}$ and} \\
                                  \nonExec{c_2}
                                \end{array}\\\\
%
                                (\mem,\reg',\stk',\ms_\stk') &
                                \arraycolsep=0pt
                                \begin{array}[t]{l}
                                  \Phi(r_1) = \retptrc \text{ and } \\
                                  \Phi(r_2) = \retptrd \text{ and } \\
                                  \Phi = (\mem, \reg, \stk, \ms_\stk) \text{ and } \\
                                  \stk = (\opc,\ms_\priv,\aaddr_\stk) :: \stk' \text{ and } \\
                                  \reg \arraycolsep=0pt
                                  \begin{array}[t]{l}
                                    \update{\pcreg}{\opc}\update{\rstk}{\stkptr{\dots,\aaddr_\stk}}\update{r_2}{0}\\
                                  \end{array}
                                \end{array}\\\\
%
                                \failed & \text{otherwise}
                              \end{cases}
\end{align*}
\lau{questions: what do we allow to be left in the registers? Everything? Only global capabilities? Whatever we do should be matched in the compilation to the target.}

@@

\lau{how should the interaction between read-only memory and linear capabilities be? That is, if a linear capability resides in read-only memory and you attempt to load it what happens? Do the load succeed with the memory cell being cleared or do the load fail (because we need to clear the cell, but we do not have the necessary permissions)? }
\begin{align*}
  \ssem{\sload{r_1}{r_2}} = & 
                              \begin{cases}
                                \updPcAddr{\update{\reg.r_1}{\var{w}}} &
                                \arraycolsep=0pt
                                \begin{array}[t]{l}
                                  \Phi = (\mem, \_, \_, \_) \tand \\
                                  \Phi(r_2) = ((\perm,\gl,\lin),\baddr,\eaddr,\aaddr) = \var{c} \tand\\
                                  \perm \in \readAllowed{} \tand \withinBounds{\var{c}} \tand \\
                                  \var{w} = \mem(\aaddr) \tand \\
                                  \nonLinear{\var{w}}
                                \end{array}\\
                                \updPcAddr{\update{\reg.r_1}{\var{w}}\update{\mem(\aaddr)}{0}} &
                                \arraycolsep=0pt
                                \begin{array}[t]{l}
                                  \Phi = (\mem, \_, \_, \_) \tand \\
                                  \Phi(r_2) = ((\perm,\gl,\lin),\baddr,\eaddr,\aaddr) = \var{c} \tand\\
                                  \perm \in \readAllowed{} \cup \writeAllowed{} \\
                                  \tand \withinBounds{\var{c}} \tand \\
                                  \var{w} = \mem(\aaddr) \tand \\
                                  \isLinear{\var{w}}
                                \end{array}\\
                                \updPcAddr{\Phi\update{\reg.r_1}{\var{w}}\update{\ms_\stk(\aaddr)}{0}} & 
                                % linear capability
                                \arraycolsep=0pt
                                \begin{array}[t]{l}
                                  \begin{array}[t]{l}
                                  \Phi = (\_, \_, \_, \ms_\stk) \tand \\
                                  \Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
                                  \perm \in \readAllowed{} \cup \writeAllowed{} \tand \\
                                  \baddr \leq \aaddr \leq \eaddr \tand \var{w} = \ms_\stk(\aaddr) \tand \\
                                  \isLinear{\var{w}}
                                \end{array}
                                \\
                                \end{array}
                                \\
                                \updPcAddr{\Phi\update{\reg.r_1}{\var{w}}} &
                                \arraycolsep=0pt
                                % non-linear capability
                                \begin{array}[t]{l}
                                  \Phi = (\_, \_, \_, \ms_\stk) \tand \\
                                  \Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
                                  \perm \in \readAllowed{} \tand \baddr \leq \aaddr \leq \eaddr \tand\\
                                  \var{w} = \ms_\stk(\aaddr) \tand \\
                                  \nonLinear{\var{w}}
                                \end{array}\\
                                \failed & \totherwise
                              \end{cases}
\end{align*}

@@

\lau{Notice, the opacity do not hide ``semantic properties'' such as permission, locality and linearity (this could be changed if it is an issue)}

@@

\textbf{gettype}\\
\lau{I changed \textbf{isptr} to an instruction that can tell what kind of word is located in a register (as we now have more kinds than integers and capabilities).}
We assume a function $\encType{} : \Word \rightarrow \ints$ that encode the different types of words (that is integers, capabilities, seals and so on) as an integer.
\lau{TODO: Write out what $\encType{}$ should satisfy}
\begin{align*}
  \tsem{\sisptr{r_1}{r_2}} = & 
                               \begin{cases}
                                 \updPcAddr{}(\Phi\updReg{r_1}{\encType{\Phi(r_2)}})
                               \end{cases}
\end{align*}

@@

\textbf{gettype}\\
\lau{I changed \textbf{isptr} to an instruction that can tell what kind of word is located in a register (as we now have more kinds than integers and capabilities).}
We assume a function $\encType{} : \Word \rightarrow \ints$ that encode the different types of words (that is integers, capabilities, seals and so on) as an integer.
\lau{TODO: Write out what $\encType{}$ should satisfy}
\begin{align*}
  \tsem{\tisptr{r_1}{r_2}} = & 
                               \begin{cases}
                                 \updPcAddr{}(\Phi\updReg{r_1}{\encType{\Phi(r_2)}})
                               \end{cases}
\end{align*}

\textbf{move}\\
\lau{Like previously, the order of updates in the second case matters for the case where $r_1 = r_2$.}
\begin{align*}
  \tsem{\tmove{r_1}{r_2}} = & 
                              \begin{cases}
                                \updPcAddr{\Phi\updReg{r_1}{\Phi(r_2)}} & \nonLinear{\Phi(r_2)}\\
                                \updPcAddr{\Phi\updReg{r_2}{0}\updReg{r_1}{\Phi(r_2)}} & \totherwise 
                              \end{cases}
\end{align*}

@@

\begin{align*}
  \tsem{\tsubseg{r_1}{r_2}{r_3}{r_4}} = &
                                     \begin{cases}
                                       \updPcAddr{\Phi\update{r_1}{\var{c}}} &
                                       \arraycolsep=0pt
                                       \begin{array}[t]{l}
                                         \Phi(r_2) = ((\perm,\gl,\lin),\baddr, \eaddr, \aaddr) \tand \\
                                         \text{for $i \in \{3,4\}$ and} \\
                                         \quad n_i = r_i \text{ or } n_i = \Phi(r_i) \tand \\
                                         \quad \text{in either case $n_i \in \ints$ and} \\
                                         \quad \baddr \leq n_3 \tand \\
                                         \quad n_4 \leq \eaddr \text{ or } n_4 = \infty \tand \\
                                         \quad \perm \neq \enter \tand \\
                                         \quad c = ((\perm,\gl,\lin),n_3,n_4,\aaddr)
                                       \end{array} \\
                                       \updPcAddr{\Phi\update{r_1}{\var{s}}} &
                                       \arraycolsep=0pt
                                       \begin{array}[t]{l}
                                         \Phi(r_2) = \seal{\perm,\phi_{\var{base}},\phi_{\var{end}}} \tand \\
                                         \text{for $i \in \{3,4\}$ and} \\
                                         \quad n_i = r_i \text{ or } n_i = \Phi(r_i) \tand \\
                                         \quad \text{in either case $n_i \in \ints$ and} \\
                                         \quad \phi_{\var{base}} \leq n_3 \tand \\
                                         \quad n_4 \leq \phi_{\var{end}} \text{ or } n_4 = \infty \tand \\
                                         \quad \perm \neq \enter \tand \\
                                         \quad c = \seal{\perm, n_3,n_4}
                                       \end{array} \\
                                       \failed & \totherwise
                                     \end{cases}
\end{align*}

@@

The program should also have access to a persistent memory which it has in the register $\rdata$ where it has a global read-write capability (or something less) (notice that this capabilty should not have execute permission). Finally the program also has access to local memory in the form of a stack. The capability for the stack resides in $\rstk$ and the capability should have local read-write-local permission (or less)
\lau{The stack capability may even need to be opaque for the sake of making the setup general (we will be handing over opaque stack capabilities).}

\lau{should the capability for the ``persistent'' memory reside along side the code?}

>>>>

\begin{document}
\section{The two capability machines}
\subsection{Domains}
Source language domains:
\[
  \begin{array}{rrcl}
   \addrbnf,\basebnf \in & \Addr & \defeq & \nats \\
    \sealbasebnf, \sigma \in & \Seal & \defeq & \nats \\
    &\Word & \defeq & \ints \uplus \SealableCaps\\
    \permbnf \in& \Perm & \defbnf & \dots \\
    &\glbnf & \defbnf & \glob \mid \local \\
    &\linbnf & \defbnf & \linear \mid \normal \\
    &\aendbnf & \in & \Addr \uplus \{\infty \} \\
    &\sealendbnf & \in & \Seal \uplus \{\infty \} \\
    &\scbnf &\defbnf & ((\permbnf,\glbnf,\linbnf),\basebnf,\aendbnf,\addrbnf) \mid \seal{\glbnf,\sealbasebnf,\sealendbnf}\\
    & & & {\sourcecolor{} \mid \stkptr{\permbnf,\basebnf,\aendbnf,\addrbnf}}\\ 
    &\cbnf & \defbnf &  \scbnf \mid \sealed{\scbnf,\sigma}{\sourcecolor{} \; \mid \retptrd \mid \retptrc}\\ 
    &\RegName & \defbnf &\pcreg \mid \rret \mid \rstk \mid \rdata \mid \dots \\
    &\Reg & \defeq & \RegName \fun \Word\\
    &\Mem & \defeq & \Addr \fun \Word \\
    &\MemSeg & \defeq & \Addr \parfun \Word \\
    {\sourcecolor{} \sstkframe \in} & {\sourcecolor{} \StkFrame} & {\sourcecolor{} \defeq} & {\sourcecolor{} \Caps \times \MemSeg \times \Addr}\\
    \src{\sstk \in}& \src{ \Stack} & \src{ \defeq} & \src{ \StkFrame^*} \\
    \Phi \in & \ExecConf & \defeq & \Mem \times \Reg {\sourcecolor{} \; \times \; \Stack \times \MemSeg }\\
    &\Conf & \defeq & \ExecConf \uplus \{\failed\} \uplus (\{\halted\} \times \Mem)
  \end{array}
\]
\lau{06-09-2017: Should seals also have the option to be linear?}
The target language domains are all the non blue parts in the above. The source language domains are the black and blue parts in the above. Further
\begin{itemize}
\item $\glbnf$ defines domain $\Global$
\item $\linbnf$ defines domain $\Linear$
\item $\scbnf$ defines domain $\SealableCaps$
\item $\cbnf$ defines domain $\Caps$
\item $\Perm$ is defined as the set of permissions in Figure~\ref{fig:perm-hier}.
\end{itemize}

@@

The source machine is also a capability machine with memory capabilities, local capabilities and sealed capabilities. Unlike the target machine, the source machine is going to have a built in stack along with special stack and return capabilities. The syntax of the source machine language is as follows:
\[
  \begin{array}{rcl}
    n & \in & \nats \\
    \src{r} &\in &  \RegName \setminus \{\rstk\}\\
    \src{\rn} &\defbnf & \src{r} \mid n \\
    \src{i} & \defbnf &  \trg{i} \mid \scall{\src{r}}{\src{r}}
  \end{array}
\]
\lau{30-08-2017: We had $\rret$ as not being one of the registers on the source machine. I could not recall why we had this, so I removed it. (I think it may have been from when we had call and return instructions?)}
There is one syntactic difference between the source language and the target language, namely the target language has an extra instruction in the $\scall{}{}$ instruction.

@@

  \caption{Permission hierarchy}
  \label{fig:perm-hier}
\end{figure}
\lau{30-08-2017: We should maybe add a picture with local/global hieararchy.}
\lau{01-09-2017: For now, I will add the linear capabilities such that a linear capability stays linear. It does, however, seem like it would be okay to have linear capabilities become non-linear (I at least don't see anything that would break down completely). One draw back would be that one could ``ruin'' the stack capability by making it non-linear.}

@@

\textbf{Call}\\
Call has three cases, a successful call, undefined and failed. For a call to be successful, it needs two sealed capabilities - a data and a code capability. The code capability must be non-executable and the stack register must contain a stack pointer. Futher the code and data capability must be local and there must be no attempt at passing on the stack capability.
\lau{30-08-2017: Add more description here.}

For the call to be undefined, the conditions necessary for a legal call must be met and there must be an attempt at passing the stack pointer to the callee or the called capabilities must be non-local \lau{30-08-2017: I have forgotten why this is an issue, but I think it is because we are crossing a security domain?}

@@

Unlike CHERI, we want sealed capabilies to be useable in a somewhat normal jump as well as the $\scall{}{}$. We cannot just use $\sjmp{}$ because it only takes one argument and we need to be able to jump to a code/data pair. To this end we introduce $\sxjmp{}{}$:
\begin{align*}
  \ssem{\sxjmp{r_1}{r_2}} = & 
                              \begin{cases}
                                \Phi'
                                \arraycolsep=0pt
                                \begin{array}[t]{l}
                                  [\reg.\pcreg \mapsto c_1] \\
                                  [\reg.\rdata \mapsto c_2]
                                \end{array} & 
                                \arraycolsep=0pt
                                \begin{array}[t]{l}
                                  \Phi(r_1) = \sealed{c_1,\sigma} \text{ and } \\
                                  \Phi(r_2) = \sealed{c_2,\sigma} \text{ and } \\
                                  \text{ for $i \in \{1,2\}$ if $\isLinear{c_i}$, then $\Phi'(r_i) = 0$} \\
                                  \text{ otherwise $\Phi'(r_i) = \Phi(r_i)$ and} \\
                                  \text{ $ \Phi(r) = \Phi'(r) \text{ for } r\in \tRegName \setminus \{r_1,r_2\}$ and} \\
                                  \nonExec{c_2}
                                \end{array}\\\\
%
                                (\mem,\reg',\stk',\ms_\stk') &
                                \arraycolsep=0pt
                                \begin{array}[t]{l}
                                  \Phi(r_1) = \retptrc \text{ and } \\
                                  \Phi(r_2) = \retptrd \text{ and } \\
                                  \Phi = (\mem, \reg, \stk, \ms_\stk) \text{ and } \\
                                  \stk = (\opc,\ms_\priv,\aaddr_\stk) :: \stk' \text{ and } \\
                                  \reg \arraycolsep=0pt
                                  \begin{array}[t]{l}
                                    \update{\pcreg}{\opc}\update{\rstk}{\stkptr{\dots,\aaddr_\stk}}\update{r_2}{0}\\
                                  \end{array}
                                \end{array}\\\\
%
                                \failed & \text{otherwise}
                              \end{cases}
\end{align*}
\lau{30-08-2017: questions: what do we allow to be left in the registers? Everything? Only global capabilities? Whatever we do should be matched in the compilation to the target.}

@@

\lau{01-09-2017: how should the interaction between read-only memory and linear capabilities be? That is, if a linear capability resides in read-only memory and you attempt to load it what happens? Do the load succeed with the memory cell being cleared or do the load fail (because we need to clear the cell, but we do not have the necessary permissions)? }
\begin{align*}
  \ssem{\sload{r_1}{r_2}} = & 
                              \begin{cases}
                                \updPcAddr{\update{\reg.r_1}{\var{w}}} &
                                \arraycolsep=0pt
                                \begin{array}[t]{l}
                                  \Phi = (\mem, \_, \_, \_) \tand \\
                                  \Phi(r_2) = ((\perm,\gl,\lin),\baddr,\eaddr,\aaddr) = \var{c} \tand\\
                                  \perm \in \readAllowed{} \tand \withinBounds{\var{c}} \tand \\
                                  \var{w} = \mem(\aaddr) \tand \\
                                  \nonLinear{\var{w}}
                                \end{array}\\
                                \updPcAddr{\update{\reg.r_1}{\var{w}}\update{\mem(\aaddr)}{0}} &
                                \arraycolsep=0pt
                                \begin{array}[t]{l}
                                  \Phi = (\mem, \_, \_, \_) \tand \\
                                  \Phi(r_2) = ((\perm,\gl,\lin),\baddr,\eaddr,\aaddr) = \var{c} \tand\\
                                  \perm \in \readAllowed{} \cup \writeAllowed{} \\
                                  \tand \withinBounds{\var{c}} \tand \\
                                  \var{w} = \mem(\aaddr) \tand \\
                                  \isLinear{\var{w}}
                                \end{array}\\
                                \updPcAddr{\Phi\update{\reg.r_1}{\var{w}}\update{\ms_\stk(\aaddr)}{0}} & 
                                % linear capability
                                \arraycolsep=0pt
                                \begin{array}[t]{l}
                                  \begin{array}[t]{l}
                                  \Phi = (\_, \_, \_, \ms_\stk) \tand \\
                                  \Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
                                  \perm \in \readAllowed{} \cup \writeAllowed{} \tand \\
                                  \baddr \leq \aaddr \leq \eaddr \tand \var{w} = \ms_\stk(\aaddr) \tand \\
                                  \isLinear{\var{w}}
                                \end{array}
                                \\
                                \end{array}
                                \\
                                \updPcAddr{\Phi\update{\reg.r_1}{\var{w}}} &
                                \arraycolsep=0pt
                                % non-linear capability
                                \begin{array}[t]{l}
                                  \Phi = (\_, \_, \_, \ms_\stk) \tand \\
                                  \Phi(r_2) = \stkptr{\perm,\baddr,\eaddr,\aaddr} \tand \\
                                  \perm \in \readAllowed{} \tand \baddr \leq \aaddr \leq \eaddr \tand\\
                                  \var{w} = \ms_\stk(\aaddr) \tand \\
                                  \nonLinear{\var{w}}
                                \end{array}\\
                                \failed & \totherwise
                              \end{cases}
\end{align*}

@@

\lau{01-09-2017: Notice, the opacity do not hide ``semantic properties'' such as permission, locality and linearity (this could be changed if it is an issue)}

@@

\textbf{gettype}\\
\lau{06-09-2017: I changed \textbf{isptr} to an instruction that can tell what kind of word is located in a register (as we now have more kinds than integers and capabilities).}
We assume a function $\encType{} : \Word \rightarrow \ints$ that encode the different types of words (that is integers, capabilities, seals and so on) as an integer.
\lau{TODO: 06-09-2017: Write out what $\encType{}$ should satisfy}
\begin{align*}
  \tsem{\sisptr{r_1}{r_2}} = & 
                               \begin{cases}
                                 \updPcAddr{}(\Phi\updReg{r_1}{\encType{\Phi(r_2)}})
                               \end{cases}
\end{align*}

@@

\textbf{gettype}\\
\lau{06-09-2017: I changed \textbf{isptr} to an instruction that can tell what kind of word is located in a register (as we now have more kinds than integers and capabilities).}
We assume a function $\encType{} : \Word \rightarrow \ints$ that encode the different types of words (that is integers, capabilities, seals and so on) as an integer.
\lau{TODO: 06-09-2017: Write out what $\encType{}$ should satisfy}
\begin{align*}
  \tsem{\tisptr{r_1}{r_2}} = & 
                               \begin{cases}
                                 \updPcAddr{}(\Phi\updReg{r_1}{\encType{\Phi(r_2)}})
                               \end{cases}
\end{align*}

\textbf{move}\\
\lau{06-09-2017: Like previously, the order of updates in the second case matters for the case where $r_1 = r_2$.}
\begin{align*}
  \tsem{\tmove{r_1}{r_2}} = & 
                              \begin{cases}
                                \updPcAddr{\Phi\updReg{r_1}{\Phi(r_2)}} & \nonLinear{\Phi(r_2)}\\
                                \updPcAddr{\Phi\updReg{r_2}{0}\updReg{r_1}{\Phi(r_2)}} & \totherwise 
                              \end{cases}
\end{align*}

@@

\begin{align*}
  \tsem{\tsubseg{r_1}{r_2}{r_3}{r_4}} = &
                                     \begin{cases}
                                       \updPcAddr{\Phi\update{r_1}{\var{c}}} &
                                       \arraycolsep=0pt
                                       \begin{array}[t]{l}
                                         \Phi(r_2) = ((\perm,\gl,\lin),\baddr, \eaddr, \aaddr) \tand \\
                                         \text{for $i \in \{3,4\}$ and} \\
                                         \quad n_i = r_i \text{ or } n_i = \Phi(r_i) \tand \\
                                         \quad \text{in either case $n_i \in \ints$ and} \\
                                         \quad \baddr \leq n_3 \tand \\
                                         \quad n_4 \leq \eaddr \text{ or } n_4 = \infty \tand \\
                                         \quad \perm \neq \enter \tand \\
                                         \quad c = ((\perm,\gl,\lin),n_3,n_4,\aaddr)
                                       \end{array} \\
                                       \updPcAddr{\Phi\update{r_1}{\var{s}}} &
                                       \arraycolsep=0pt
                                       \begin{array}[t]{l}
                                         \Phi(r_2) = \seal{\perm,\phi_{\var{base}},\phi_{\var{end}}} \tand \\
                                         \text{for $i \in \{3,4\}$ and} \\
                                         \quad n_i = r_i \text{ or } n_i = \Phi(r_i) \tand \\
                                         \quad \text{in either case $n_i \in \ints$ and} \\
                                         \quad \phi_{\var{base}} \leq n_3 \tand \\
                                         \quad n_4 \leq \phi_{\var{end}} \text{ or } n_4 = \infty \tand \\
                                         \quad \perm \neq \enter \tand \\
                                         \quad c = \seal{\perm, n_3,n_4}
                                       \end{array} \\
                                       \failed & \totherwise
                                     \end{cases}
\end{align*}
\dominique{31-8-2017: Why do we allow applying subseg to a sealed capability?}
\lau{6-9-2017: We do not. We allow subseg on the seals that are used to seal capabilities.}

@@

The program should also have access to a persistent memory which it has in the register $\rdata$ where it has a global read-write capability (or something less) (notice that this capabilty should not have execute permission). Finally the program also has access to local memory in the form of a stack. The capability for the stack resides in $\rstk$ and the capability should have local read-write-local permission (or less)
\lau{30-08-2017: The stack capability may even need to be opaque for the sake of making the setup general (we will be handing over opaque stack capabilities).}

\lau{30-08-2017: should the capability for the ``persistent'' memory reside along side the code?}
